---
title: "Assignment 6, task 3: lizard length vs species"
author: "Elmera Azadpour"
date: "11/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(effsize)
library(broom)
```
#### Introduction:
Utilizing data from the lizard pitfall trap data (LTER-II, LTER-III) ver 36: Environmental Data Initiative (https://doi.org/10.6073/pasta/ff37b4cace16a9943575f3fd7067064e), I compared total lizard length (mm) of two species: CNUN (*Cnemidophorus uniparens*, the Desert Grassland lizard) and CNTE (*Cnemidophorus tessalatus*, the Colorado Checkered lizard). To examine a significant difference in total length difference in mean total length between female lizards for the two species, I conducted a Welch’s two-sample t-test and Cohen's d effect size analysis. 


```{r, error=FALSE, warning=FALSE, message=FALSE, echo = FALSE, include = FALSE}
lizards <- read_csv(here("data", "lizards.csv"))
```


```{r, error=FALSE, warning=FALSE, message=FALSE, echo = FALSE, include=FALSE}
lizards_clean <- lizards %>%
  select(spp, sex, total_length) %>% 
  filter(sex == "F") %>% 
  filter(spp %in% c("CNUN", "CNTE"))
```


```{r, error=FALSE, warning=FALSE, message=FALSE, echo = FALSE, fig.cap = "**Figure 1**.  Boxplot comparing total lengths of female lizards for CNUN (*Cnemidophorus uniparens*, the Desert Grassland lizard) and CNTE (*Cnemidophorus tessalatus*, the Colorado Checkered lizard),"}
####  Data visualization comparing total lengths of female lizards for the two species: CNUN and CNTE
ggplot(data = lizards_clean,
       aes(x = spp, 
           y = total_length)) + 
  geom_boxplot(aes(fill = spp)) +
  theme_minimal() +
  theme(legend.position = "none") + 
  labs(x = "Species", y = "Total length (mm)")
```


```{r, error=FALSE, warning=FALSE, message=FALSE, echo = FALSE}
####  Summary table that contains the mean and SD for total length, and sample size, for each group (female CNUN and CNTE lizards)
lizards_clean_table <- lizards_clean %>% 
 group_by(spp) %>% 
  summarise(mean_length = mean(total_length, na.rm=T),
            sd_length = sd(total_length, na.rm=T),
            n = n())

lizards_clean_table %>%  
  kable(col.names = c("Species",
                      "Mean length (mm)",
                      "Standard deviation",
                      "Sample size (n)"), 
        caption = "**Table 1**: summary statistics of female CNUN and CNTE lizards") %>% 
  kable_styling(full_width = FALSE) 
```


```{r, error=FALSE, warning=FALSE, message=FALSE, echo = FALSE, include = FALSE}
# Histograms
ggplot(data = lizards_clean, aes(x = total_length)) +
  geom_histogram(bins = 15) +
  facet_wrap(~spp, scales = "free")

# QQ Plots
ggplot(data= lizards_clean, aes(sample = total_length)) +
  geom_qq() +
  facet_wrap(~spp)
```


```{r, error=FALSE, warning=FALSE, message=FALSE, echo = FALSE, include = FALSE}
# create vectors 
lizard_length_cnte <- lizards_clean %>% 
  select(c("spp", "total_length")) %>% 
  filter(spp == "CNTE") %>% 
  pull(total_length)

lizard_length_cnun <- lizards_clean %>% 
  select(c("spp", "total_length")) %>% 
  filter(spp == "CNUN") %>% 
  pull(total_length)

ttest <-t.test(lizard_length_cnte, lizard_length_cnun)
ttest_tidy <- tidy(ttest)


# cohens d/ effect size
cohen_test <- cohen.d(lizard_length_cnte, lizard_length_cnun, na.rm = TRUE)
```

#### Summary:
+ This analysis indicates CNTE (*Cnemidophorus tessalatus*, the Colorado Checkered lizard) had a larger mean (244.90 ± 47.339, *n* = 28; mean ± 1 standard deviation), compared to CNUN (*Cnemidophorus uniparens*, the Desert Grassland lizard) which had a smaller mean (147.60 ± 34.55, *n* = 47; mean ± 1 standard deviation). The actual difference in means from CNTE and CNUN lizards is 97.2929. The outcome of the two sample t-test indicated there is not a strong change (*p* < 0.001) of randomly selecting two samples from populations with the same that are this difference by change. In sum, the difference in means is significant (Welch’s two-sample t-test: t(`r round(ttest_tidy$parameter,2)`) = `r round(ttest_tidy$statistic,2)`, *p* < 0.001), and the effect size is large (Cohen's *d* = `r round(cohen_test$estimate,2)`).


#### Citations:
+ Data source: Lightfoot, D. 2017. Lizard pitfall trap data (LTER-II, LTER-III) ver 36. Environmental Data Initiative. https://doi.org/10.6073/pasta/ff37b4cace16a9943575f3fd7067064e 

+ Metadata: https://portal.edirepository.org/nis/metadataviewer?packageid=knb-lter-jrn.210007001.36
